---
title: "River - Test Setup"
author: "River Mckegney"
date: "`r Sys.Date()`"
output: html_document
---

## Setup, and Packages

```{r setup, include=FALSE}
# Setup
knitr::opts_chunk$set(echo = TRUE)
```

Code Packages:
```{r packages}
# Load packages
library(gridExtra) # For using grid.table to save tables as images
library(flextable) # Alternative method to save df as image
library(tidyverse)
library(zoo)
library(lme4)
library(lmerTest)
library(rlang)
library(nortest) # For ad.test
library(car) # For homogeneity of variance with leveneTest
library(rcompanion) # For scheirerRayHare test
library(pastecs) # For stat.desc
library(ggpubr) # For assumption plots
library(conover.test) # For the conover.Iman() test function
library(DescTools) # Contain alternative ConoverTest function
library(rstatix) # tidyverse adapted stat tests and mshapiro_test() functions
library(DHARMa) # For multi-level lm models and shapiro test
library(patchwork) # For combining ggplots
```

## Data Entry: Read-in data

```{r}
# Read and call data into df

# Primary samples df
ABL90 <- read.csv("Raw_Data/ABL90_Raw.csv")

# River's Parturition metadata
partuition_subcat <- read.csv("Raw_Data/River's_Rockfish_Metadata_Parturition_V7.csv")
```


## Data Filtering

```{r}
# Data sifting: ABL90 dataset

# Step 1: Remove missing info
ABL_set1 <- ABL90 %>%
  filter(Patient.ID_edited != "") 
  

# Step 2: Separate Patient.ID into columns of sample types: blood plasma and instant freeze plasma 
ABL_set2 <- separate(ABL_set1, 'Patient.ID_edited', into = c("Patient.ID_edited", "Sample_Type"), sep = ",")  


# Step 3:
# Convert Patient.ID_edited column data to character type
ABL_set2$Patient.ID_edited <- as.character(ABL_set2$Patient.ID_edited) 


```


```{r}
# Step 4 & 5: Sussing out specific sample errors  

# Remove insufficient samples
ABL_set3 <- ABL_set2 %>%
  filter(!is.na("Type")) %>%
  filter(!str_detect(Errors.detected.during.measurement, "Insufficient sample"))

# Remove inhomogeneous samples
ABL_set3 <- ABL_set3 %>%
  filter(!is.na("Type")) %>%
  filter(!str_detect(Errors.detected.during.measurement, "Inhomogeneous sample"))

# Step 6: Filter for only blood samples
ABL_b_samp <- ABL_set3 %>%
  filter(Sample_Type == "b")

```


## Merge Datasets:

Join ABL90 df with parturition_subcat df: Making ABL_merged
```{r}
# Rename columns to match: Change 'ID' in parturition_subcat to 'Patient.ID_edited' so it is ready to join with ABL90 df

# Please note the 'Treatment' col in parturition_subcat does not match with the finalized 'Ambient.Or.OAH' col in metadata_atresia_guide or ABL90_merged df

# Rename 'ID' col to 'patient.ID_edited'
partuition_subcat <- partuition_subcat %>%
  rename(Patient.ID_edited = ID)

```

```{r}
# Connect main ABL90 dataset with my (River's) sub categorized parturition metadata 
# ABL_merged <- partuition_subcat %>%
# left_join(ABL_b_samp, by = "Patient.ID_edited")

ABL_merged <- ABL_b_samp %>%
inner_join(partuition_subcat, by = "Patient.ID_edited")
```

Rename Columns Neatly:
```{r}
# Rename Treatment & Parturition Outcome
ABL_merged <- ABL_merged %>%
  rename(Ambient.Or.OAH = Consensus_General_Treatment,
         Pregnant.Or.Atresia = Consensus_Atresia_Or_Pregnant)
         
```


## Remove Samples: Mortalities, NA's, or Missing Info, and Replicates

Remove Replicate ID's
```{r}
# Cross Check all Columns: Cross validation to suss correct replicate row
check_params <- ABL_merged %>%
  select(Patient.ID_edited, Time, Sample.., Measuring.Mode, Ambient.Or.OAH, Pregnant.Or.Atresia, pH, Glu..mmol.L., Hct...., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L., pCO2..mmHg.) %>%
  filter(Patient.ID_edited == "9782D")
print(check_params)

check_params <- ABL_merged %>%
  select(Patient.ID_edited, Time, Sample.., Measuring.Mode, Ambient.Or.OAH, Pregnant.Or.Atresia, pH, Glu..mmol.L., Hct...., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L., pCO2..mmHg.) %>%
  filter(Patient.ID_edited == "9783D")
print(check_params)

# Removed replicates: 9782D (2x) and 9783D (2x)
ABL_merged <- ABL_merged %>%
  filter(Sample.. != "863",
         Sample.. != "934")

```


Remove Motalities and No info IDs and assign analysis ready data-frames:
```{r}

# New df with Moralities removed: Note none of these samples made it into ABL90 df anyway, so looks like they are already filtered out.
 Live_Samples <- ABL_merged %>%
  filter(Patient.ID_edited != "9780C", # Mortality
         Patient.ID_edited != "777AE", # Mortality
         Patient.ID_edited != "777CA") # Mortality after parturition


# New df with mortality and 'No info' Id's removed
 Primary_Samples <- Live_Samples %>%
   filter(Patient.ID_edited != "777A0", # No info
           Patient.ID_edited != "9782F", # No info
           Patient.ID_edited != "777B3", # No info
           Patient.ID_edited != "777AA", # No info
           Patient.ID_edited != "777DE", # No info
           Patient.ID_edited != "777CE", # No info
           Patient.ID_edited != "777A6") # No info
 
 
# New df of Only Ambient Treatment: For testing parturition success
 Ambient_Samples <- Primary_Samples %>%
   filter(Ambient.Or.OAH == "Ambient")
 
# New df of Fecundity Samples: For Background Correlative Evidence To Compare with Brood Condition Categorization
 Fecundity_Samples <- Primary_Samples %>%
   filter(Fecundity_Class != "NA")
 
```



Change Data Types:
```{r}
# Change Continuous Data to type to numeric, double, or factor

# Change data type of ions to factor or numeric
#glimpse(General_Samples)
#glimpse(Ambient_Samples)
 
Primary_Samples <- Primary_Samples %>%
  mutate(across(c(pH, Hct...., Glu..mmol.L., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L.), as.numeric))

Ambient_Samples <- Ambient_Samples %>%
  mutate(across(c(pH, Hct...., Glu..mmol.L., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L.), as.numeric))

Fecundity_Samples <- Fecundity_Samples %>%
  mutate(across(c(Fecundity_Or_Brood_Count, pH, Hct...., Glu..mmol.L., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L.), as.numeric))

# glimpse(Primary_Samples)
# glimpse(Ambient_Samples)
# glimpse(Fecundity_Samples)
```


Order Factor Levels
```{r}
# Primary data

# Arrange the order of parturition categories for visualizations & tests
Primary_Samples$Ambient.Or.OAH <- ordered(Primary_Samples$Ambient.Or.OAH, levels = c("Ambient", "OAH"))

Primary_Samples$Pregnant.Or.Atresia <- ordered(Primary_Samples$Pregnant.Or.Atresia, levels = c("Post Parturition", "Atresia"))

```


## Sample Size: n
```{r}
# Sample Size

# Primary data: 
primary_sample_table <- Primary_Samples %>%
  group_by(Ambient.Or.OAH, Pregnant.Or.Atresia) %>%
  summarise(n_size = n_distinct(Patient.ID_edited), .groups = "drop")
print(primary_sample_table)

 pdf("data-images/primary_sample_table.pdf")
 grid.table(primary_sample_table)
 dev.off()
 
 png("data-images/primary_sample_table.png")
 grid.table(primary_sample_table)
 dev.off()

```


Sample Size Barplot:
```{r}
# View Sample Distributions

# Sample Size barplot

# Primary Samples
primary_sample_barplot <- Primary_Samples %>%
  group_by(Pregnant.Or.Atresia, Ambient.Or.OAH) %>% 
  summarise(n_size = n_distinct(Patient.ID_edited), .groups = "drop") %>%
  ggplot(aes(x = Pregnant.Or.Atresia, y = n_size, fill = Pregnant.Or.Atresia)) +
  geom_bar(stat = "identity" ) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 15.5), breaks = c(0, 3, 6, 9, 12, 15)) +
  facet_grid(~ Ambient.Or.OAH) +
  labs(title = "Primary Sample Size",
        x = "Parturition Status & Treatment Type",
        y = "Sample Size (n)") +
  guides(fill = guide_legend((title = "Reproductive State"))) +
  theme_classic(base_size = 10)

print(primary_sample_barplot)

ggsave(filename = "data-images/.png", plot = primary_sample_barplot, device = "png")
ggsave(filename = "data-images/primary_sample_barplot.pdf", plot = primary_sample_barplot, device = "pdf")

```


# pH

pH Boxplot
```{r}
# pH boxplot

# Primary samples
pH_primary_boxplot <- ggplot(data = Primary_Samples, aes(x = Pregnant.Or.Atresia, y = pH, fill = Ambient.Or.OAH)) +
  geom_boxplot(aes(x = Pregnant.Or.Atresia, y = pH, fill = Ambient.Or.OAH)) +
  geom_point(aes(color = Ambient.Or.OAH), position = position_dodge(width = 0.75), alpha = 0.3, color = "black") +
  labs(title = "Primary pH", x = "Parturition Type", y = "pH") +
  guides(fill = guide_legend((title = "Treatment Type"))) + 
  scale_fill_manual(values = c("Ambient" = "#189bff", "OAH" = "#598c78")) +
  theme_classic() +
  theme(panel.background = element_rect(fill = "white"),
        plot.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white", color = "black"),
        title = element_text(size = 12),
        plot.title = element_text(size = 30, hjust = 0.5),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 20),
        axis.title.x = element_text(size = 20),
        axis.title.y = element_text(size = 20))

print(pH_primary_boxplot)

ggsave(filename = "data-images/pH_primary_boxplot.pdf", plot = pH_primary_boxplot, device = "pdf")
ggsave(filename = "data-images/pH_primary_boxplot.png", plot = pH_primary_boxplot, device = "png")

```


pH Parametric Assessment
Data Distribution
```{r}
# pH
# Parametric Assumptions

# Data Distribution

# Frequency Histograms

hist(Primary_Samples$pH)


# Density vs normality curve histograms

Primary_Samples %>%
  ggplot(aes(x = pH)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm, args = c(mean = mean(Primary_Samples$pH), sd = sd(Primary_Samples$pH)), aes(col = "Normal Distribution"), show.legend = TRUE) +
  geom_density(aes(col = "Sample Density"), show.legend = TRUE) +
  labs(title = "pH Primary Sample Distribution") +
  scale_color_manual("Distribution", values = c("Normal Distribution" = "green", "Sample Density" = "black"))

```

pH Parametric Assessment

Model Design
```{r}

# Install flexplot
# devtools::install_github("dustinfife/flexplot", ref="development")
require(flexplot)

# Best guess method: Treatment fixed, 
# lm model of primary data
pH_primary_lmer <- lmer(pH ~ Ambient.Or.OAH * Pregnant.Or.Atresia + (1|Pregnant.Or.Atresia), data = Primary_Samples)

# Second best guess method:
pH_primary_lmer2 <- lmer(pH ~ Ambient.Or.OAH * Pregnant.Or.Atresia + (1|Pregnant.Or.Atresia), data = Primary_Samples)

# Third option
pH_primary_lmer3 <- lmer(pH ~ Pregnant.Or.Atresia * Ambient.Or.OAH + (1|Pregnant.Or.Atresia) + (1|Ambient.Or.OAH), data = Primary_Samples)

# What's up with that weird intersept (why does clusters either start or finish at same location, they should both be fixed right?)

visualize(pH_primary_lmer, plot = "model", sample = 50)
visualize(pH_primary_lmer2, plot = "model", sample = 50)
visualize(pH_primary_lmer3, plot = "model", sample = 50)

```



Normality Residual QQplot
```{r}
# pH
# Parametric Assumptions:

# Assessing Normality

# Summary of model
summary(pH_primary_lmer)

summary(pH_primary_lmer2)

summary(pH_primary_lmer3)

# Suss residuals of model
pH_primary_res_lmer <- residuals(pH_primary_lmer)

pH_primary_res_lmer2 <- residuals(pH_primary_lmer2)

pH_primary_res_lmer3 <- residuals(pH_primary_lmer3)

# Shapiro test
shapiro.test(pH_primary_res_lmer)

shapiro.test(pH_primary_res_lmer2)

shapiro.test(pH_primary_res_lmer3)

# pH simulated residuals plot
pH_primary_simres_plot <- simulateResiduals(fittedModel = pH_primary_lmer)
plot(pH_primary_simres_plot)

pH_primary_simres_plot2 <- simulateResiduals(fittedModel = pH_primary_lmer2)
plot(pH_primary_simres_plot2)

pH_primary_simres_plot3 <- simulateResiduals(fittedModel = pH_primary_lmer3)
plot(pH_primary_simres_plot3)

# Qqplot of model
# pH_primary_res_qqplot <- ggqqplot(residuals(pH_primary_lmer)) +
# labs(title = "Primary pH Interactive Residual QQplot",
#      subtitle = "lmer(pH ~ Pregnant.Or.Atresia + (1|Ambient.Or.OAH), data = Primary_Samples)",
#                                x = "pH Theoretical Quantiles (Predicted values)",
#                                y = "pH Sample Quantiles") +
#   theme(plot.title = element_text(hjust = 0.5),
#         plot.subtitle = element_text(hjust = 0.5))
# print(pH_primary_res_int_qqplot)
```


pH Parametric Assessment
Testing Scedasticity:
```{r}
# pH

# Primary samples

# test with lmer
bartlett.test(pH_primary_res_lmer ~ Primary_Samples$Ambient.Or.OAH)

bartlett.test(pH_primary_res_lmer ~ Primary_Samples$Pregnant.Or.Atresia)

bartlett.test(pH ~ Ambient.Or.OAH, Primary_Samples)
bartlett.test(pH ~ Pregnant.Or.Atresia, Primary_Samples) # Significant

leveneTest(pH ~ Pregnant.Or.Atresia, Primary_Samples) # Slightly not significant
```


pH Parametric Assumptions Assessment: 

- Primary Data normally distributed?

- Primary Data equal variance across factors?


# pH Stat Tests
```{r}
# pH Analysis
# ANOVA Tests

# Testing primary factor differences: Treatment & Parturition Outcome

# Two-Way Interactive ANOVA
pH_primary_aov_int <- aov(pH ~ Pregnant.Or.Atresia * Ambient.Or.OAH, data = Primary_Samples)
summary(pH_primary_aov_int)

# If interaction not significant (p-value > 0.05), use additive model

pH_primary_aov_additive <- aov(pH ~ Pregnant.Or.Atresia + Ambient.Or.OAH, data = Primary_Samples)
summary(pH_primary_aov_additive)

# Post-Hoc Test: If parametric Tukeys, if nonparametric Aligned Ranked Transformation ANOVA (ART)

# How do I interpret that...
pH_tukey_test <- TukeyHSD(pH_primary_aov_int)
print(pH_tukey_test)

#pH_primary_aov_int %>% tukey_hsd()
```
pH Results:

Interaction significant?

- Treatment (alone) OAH vs Ambient: (p adj = )

- Pregnancy (alone) Post-parturition vs Atresia: (p adj = )

- Post parturition vs OAH: (p adj = )

- Post Parturition vs Ambient: (p adj = )

- Atresia vs Ambient: (p adj = )

- Atresia vs OAH: (p adj = )


