---
title: "SU25_River_Rockfish_Proj_Density-Plots"
author: "River Mckegney"
date: "`r Sys.Date()`"
output: html_document
---


## Setup, and Packages

```{r setup, include=FALSE}
# Setup
knitr::opts_chunk$set(echo = TRUE)
```

Code Packages:
```{r packages}
# Load packages
library(gridExtra) # For using grid.table to save tables as images
library(flextable) # Alternative method to save df as image
library(tidyverse)
library(zoo)
library(lme4)
library(lmerTest)
library(rlang)
library(nortest) # For ad.test
library(car) # For homogeneity of variance with leveneTest
library(rcompanion) # For scheirerRayHare test
library(pastecs) # For stat.desc
library(ggpubr) # For assumption plots
library(conover.test) # For the conover.Iman() test function
library(DescTools) # Contain alternative ConoverTest function
library(rstatix) # tidyverse adapted stat tests and mshapiro_test() functions
library(DHARMa) # For multi-level lm models and shapiro test
library(patchwork) # For combining ggplots
```

## Data Entry: Read-in data

```{r}
# Read and call data into df

# Primary samples df
ABL90 <- read.csv("Raw_Data/ABL90_Raw.csv")

# River's Parturition metadata
partuition_subcat <- read.csv("Raw_Data/River's_Rockfish_Metadata_Parturition_V7.csv")
```


## Data Filtering

```{r}
# Data sifting: ABL90 dataset

# Step 1: Remove missing info
ABL_set1 <- ABL90 %>%
  filter(Patient.ID_edited != "") 
  

# Step 2: Separate Patient.ID into columns of sample types: blood plasma and instant freeze plasma 
ABL_set2 <- separate(ABL_set1, 'Patient.ID_edited', into = c("Patient.ID_edited", "Sample_Type"), sep = ",")  


# Step 3:
# Convert Patient.ID_edited column data to character type
ABL_set2$Patient.ID_edited <- as.character(ABL_set2$Patient.ID_edited) 


```


```{r}
# Step 4 & 5: Sussing out specific sample errors  

# Remove insufficient samples
ABL_set3 <- ABL_set2 %>%
  filter(!is.na("Type")) %>%
  filter(!str_detect(Errors.detected.during.measurement, "Insufficient sample"))

# Remove inhomogeneous samples
ABL_set3 <- ABL_set3 %>%
  filter(!is.na("Type")) %>%
  filter(!str_detect(Errors.detected.during.measurement, "Inhomogeneous sample"))

# Step 6: Filter for only blood samples
ABL_b_samp <- ABL_set3 %>%
  filter(Sample_Type == "b")

```


## Merge Datasets:

Join ABL90 df with parturition_subcat df: Making ABL_merged
```{r}
# Rename columns to match: Change 'ID' in parturition_subcat to 'Patient.ID_edited' so it is ready to join with ABL90 df

# Please note the 'Treatment' col in parturition_subcat does not match with the finalized 'Ambient.Or.OAH' col in metadata_atresia_guide or ABL90_merged df

# Rename 'ID' col to 'patient.ID_edited'
partuition_subcat <- partuition_subcat %>%
  rename(Patient.ID_edited = ID)

```

```{r}
# Connect main ABL90 dataset with my (River's) sub categorized parturition metadata 
# ABL_merged <- partuition_subcat %>%
# left_join(ABL_b_samp, by = "Patient.ID_edited")

ABL_merged <- ABL_b_samp %>%
inner_join(partuition_subcat, by = "Patient.ID_edited")
```

Rename Columns Neatly:
```{r}
# Rename Treatment & Parturition Outcome
ABL_merged <- ABL_merged %>%
  rename(Ambient.Or.OAH = Consensus_General_Treatment,
         Pregnant.Or.Atresia = Consensus_Atresia_Or_Pregnant)
         
```


## Remove Samples: Mortalities, NA's, or Missing Info, and Replicates

Remove Replicate ID's
```{r}
# Cross Check all Columns: Cross validation to suss correct replicate row
check_params <- ABL_merged %>%
  select(Patient.ID_edited, Time, Sample.., Measuring.Mode, Ambient.Or.OAH, Pregnant.Or.Atresia, pH, Glu..mmol.L., Hct...., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L., pCO2..mmHg.) %>%
  filter(Patient.ID_edited == "9782D")
print(check_params)

check_params <- ABL_merged %>%
  select(Patient.ID_edited, Time, Sample.., Measuring.Mode, Ambient.Or.OAH, Pregnant.Or.Atresia, pH, Glu..mmol.L., Hct...., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L., pCO2..mmHg.) %>%
  filter(Patient.ID_edited == "9783D")
print(check_params)

# Removed replicates: 9782D (2x) and 9783D (2x)
ABL_merged <- ABL_merged %>%
  filter(Sample.. != "863",
         Sample.. != "934")

```


Remove Motalities and No info IDs and assign analysis ready data-frames:
```{r}

# New df with Moralities removed: Note none of these samples made it into ABL90 df anyway, so looks like they are already filtered out.
 Live_Samples <- ABL_merged %>%
  filter(Patient.ID_edited != "9780C", # Mortality
         Patient.ID_edited != "777AE", # Mortality
         Patient.ID_edited != "777CA") # Mortality after parturition


# New df with mortality and 'No info' Id's removed
 Primary_Samples <- Live_Samples %>%
   filter(Patient.ID_edited != "777A0", # No info
           Patient.ID_edited != "9782F", # No info
           Patient.ID_edited != "777B3", # No info
           Patient.ID_edited != "777AA", # No info
           Patient.ID_edited != "777DE", # No info
           Patient.ID_edited != "777CE", # No info
           Patient.ID_edited != "777A6") # No info
 
 
# New df of Only Ambient Treatment: For testing parturition success
 Ambient_Samples <- Primary_Samples %>%
   filter(Ambient.Or.OAH == "Ambient")
 
# New df of Fecundity Samples: For Background Correlative Evidence To Compare with Brood Condition Categorization
 Fecundity_Samples <- Primary_Samples %>%
   filter(Fecundity_Class != "NA")
 
```



Change Data Types:
```{r}
# Change Continuous Data to type to numeric, double, or factor

# Change data type of ions to factor or numeric
#glimpse(General_Samples)
#glimpse(Ambient_Samples)
 
Primary_Samples <- Primary_Samples %>%
  mutate(across(c(pH, Hct...., Glu..mmol.L., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L.), as.numeric))

Ambient_Samples <- Ambient_Samples %>%
  mutate(across(c(pH, Hct...., Glu..mmol.L., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L.), as.numeric))

Fecundity_Samples <- Fecundity_Samples %>%
  mutate(across(c(Fecundity_Or_Brood_Count, pH, Hct...., Glu..mmol.L., Na...mmol.L., Cl...mmol.L., K...mmol.L., Ca....mmol.L.), as.numeric))

# glimpse(Primary_Samples)
# glimpse(Ambient_Samples)
# glimpse(Fecundity_Samples)
```


Order Factor Levels
```{r}
# Primary data

# Arrange the order of parturition categories for visualizations & tests
Primary_Samples$Ambient.Or.OAH <- ordered(Primary_Samples$Ambient.Or.OAH, levels = c("Ambient", "OAH"))

Primary_Samples$Pregnant.Or.Atresia <- ordered(Primary_Samples$Pregnant.Or.Atresia, levels = c("Post Parturition", "Atresia"))

```


## Ambient Density Plots
```{r}
# pH
ggplot(Ambient_Samples, aes(x = pH)) +
  geom_density(aes(x = pH)) +
  facet_wrap(~ Consensus_Brood_Condition)

# Hct
ggplot(Ambient_Samples, aes(x = Hct....)) +
  geom_density(aes(x = Hct....)) +
  facet_wrap(~ Consensus_Brood_Condition)

# Glucose
ggplot(Ambient_Samples, aes(x = Glu..mmol.L.)) +
  geom_density(aes(x = Glu..mmol.L.)) +
  facet_wrap(~ Consensus_Brood_Condition)

# Na+
ggplot(Ambient_Samples, aes(x = Na...mmol.L.)) +
  geom_density(aes(x = Na...mmol.L.)) +
  facet_wrap(~ Consensus_Brood_Condition)

# Cl-
ggplot(Ambient_Samples, aes(x = Cl...mmol.L.)) +
  geom_density(aes(x = Cl...mmol.L.)) +
  facet_wrap(~ Consensus_Brood_Condition)

# K+
ggplot(Ambient_Samples, aes(x = K...mmol.L.)) +
  geom_density(aes(x = K...mmol.L.)) +
  facet_wrap(~ Consensus_Brood_Condition)

# Ca+2
ggplot(Ambient_Samples, aes(x = Ca....mmol.L.)) +
  geom_density(aes(x = Ca....mmol.L.)) +
  facet_wrap(~ Consensus_Brood_Condition)

```


